IIIDYT at IEST 2018: Implicit Emotion Classification With Deep Contextualized Word Representations

### 2018-09-01 paper/paper.tex
[paper] Add acknowledgements, wrap lines, small fixes

\begin{document}
\maketitle
\begin{abstract}
In this paper we describe our system designed for the WASSA 2018 Implicit Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 26 teams with a test macro F1 score of $0.710$. The system is composed of a single pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory Network BiLSTM for enriching word representations with context, a max-pooling operation for creating sentence representations from said word vectors, and a Dense Layer for projecting the sentence representations into label space. Our official submission was obtained by ensembling 6 of these models initialized with different random seeds.
\end{abstract}

@@

    Furthermore, we tokenized the text using a variation of the \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}} script, a Python port of the original \texttt{Twokenize.java} \cite{gimpel-EtAl:2011:ACL-HLT2011}. Concretely, we created an emoji-aware version of it by incorporating knowledge from an emoji database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}} which we slightly modified for avoiding conflict with emoji sharing unicode codes with common glyphs used in Twitter,\footnote{For example, the hashtag emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3}, which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and for making it compatible with Python 3.
% \begin{itemize}
%     \item briefly describe previous Semeval papers~\cite{baziotis2018ntua,
%         duppada2018seernet, abdou2018affecthor} (Should we do this despite not
%         having based our work in any of them?)
%     \item mention previous Wassa papers
%     \item talk about some sentence classification tasks? Works that fall into
%         the broad category of mapping sentences to labels.
%     \item Talk about elmo~\cite{peters2018deep}
% \end{itemize}

@@

\textbf{Dimensionalities}: By default the ELMo layer outputs a 2048-dim vector, which we then feed to a BiLSTM resulting in a 4096-dim vector for each word of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read the input in opposite directions and whose outputs are concatenated at the sequence dimension. This results in a vector double the dimension of the input for each time step.}. After max-pooling the BiLSTM output over the sequence dimension we have a single 4096-dim vector corresponding to the tweet representation. This representation is finally fed to a single-layer fully-connected network with input size 4096, 512 hidden units, output size 6, and ReLU nonlinearities after the input and hidden layers. The output of the dense layer is a 6-dim logit vector for each example.

@@

\textbf{Optimization}: We optimized the model with Adam \cite{DBLP:journals/corr/KingmaB14}, using a slanted triangular learning rate schedule \cite{howard2018universal}, with default hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate $\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained by multiplying the number of epochs ($10$) times the total number of batches, which for the training dataset corresponds to $2396$ batches of $64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.

@@

\textbf{Model Selection}: To choose the best hyperparameter configuration we measured the classification accuracy on the validation set (trial set).

@@

    \caption{Ablation study results.}
    \vspace{-0.4cm}
    \caption*{
        \footnotesize Accuracies were obtained from the
        validation dataset. Each model was trained with the same random seed and
        hyperparameters, save for the one listed. ``No emoji'' is the same model
        trained on the training dataset with no emoji, ``No ELMo'' corresponds
        to having switched the ELMo word encoding layer with a simple
        pre-trained GloVe embedding lookup table, and ``Concat Pooling''
        obtained sentence representations by using the pooling method described
        by \citet{howard2018universal}. ``LSTM hidden'' corresponds to the the
        hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the
        part-of-speech embeddings, and ``SGD optim lr'' to the learning rate
        used while optimizing with the schedule described by \citet{conneau2017supervised}.}

@@

Finally, not all emoji are beneficial for this task. When removing \texttt{sweat\_smile} and \texttt{confused} accuracy increased, probably because they represent emotions other than the ones being predicted.
%From observing Table \ref{table:emoji_label} it seems likely that tweets containing \texttt{mask} would be highly predictive of the \texttt{disgust} class, since more than $92\%$ of the tweets containing such emoji belong to that class, however it less clear why \texttt{rage}, and \texttt{cry} contribute similarly, despite being more evenly distributed among classes.

%This means that they are highly representative of the the class the tweet belongs to, and would be good choices for discriminating new tweets in case we wanted to expand the dataset.

%For example, by observing Table \ref{table:emoji_label} we can infer that  \texttt{mask}, \texttt{rage}, and \texttt{cry}   \texttt{disgust}, \texttt{rage}, and \texttt{sad} classes respectively.

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}

In this paper we describe our system designed for the WASSA 2018 Implicit
Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 30 teams
with a test macro F1 score of $0.710$. The system is composed of a single
pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory
Network BiLSTM for enriching word representations with context, a max-pooling
operation for creating sentence representations from them, and a
Dense Layer for projecting the sentence representations into label space. Our
official submission was obtained by ensembling 6 of these models initialized
with different random seeds.

\end{abstract}

@@

    Furthermore, we tokenized the text using a variation of the
    \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}}
    script, a Python port of the original \texttt{Twokenize.java}
    \cite{gimpel-EtAl:2011:ACL-HLT2011}. Concretely, we created an emoji-aware
    version of it by incorporating knowledge from an emoji
    database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
    which we slightly modified for avoiding conflict with emoji sharing unicode
    codes with common glyphs used in Twitter,\footnote{For example, the hashtag
    emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3},
which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and
for making it compatible with Python 3.

% \begin{itemize}
%     \item briefly describe previous Semeval papers~\cite{baziotis2018ntua,
%         duppada2018seernet, abdou2018affecthor} (Should we do this despite not
%         having based our work in any of them?)
%     \item mention previous Wassa papers
%     \item talk about some sentence classification tasks? Works that fall into
%         the broad category of mapping sentences to labels.
%     \item Talk about elmo~\cite{peters2018deep}
% \end{itemize}

@@

\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dim vector,
which we then feed to a BiLSTM with output size 2048, resulting in a 4096-dim
vector when concatenating forward and backward directions for each word
of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read
the input in opposite directions and whose outputs are concatenated at the
hidden dimension. This results in a vector double the dimension of the input
for each time step.}. After max-pooling the BiLSTM output over the sequence
dimension, we obtain a single 4096-dim vector corresponding to the tweet
representation. This representation is finally fed to a single-layer
fully-connected network with input size 4096, 512 hidden units, output size 6,
and a ReLU nonlinearity after the hidden layer. The output of the
dense layer is a 6-dim logit vector for each input example.

@@

\textbf{Optimization}: We optimized the model with Adam
\cite{DBLP:journals/corr/KingmaB14}, using default hyperparameters
($\beta_1=0.9$, $\beta_2=0.999$, $\epsilon=10^{-8}$), following a slanted
triangular learning rate schedule \cite{howard2018universal}, also with default
hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate
$\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained
    by multiplying the number of epochs ($10$) times the total number of
    batches, which for the training dataset corresponds to $2396$ batches of
$64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.

@@

\textbf{Model Selection}: To choose the best hyperparameter configuration we
measured the classification accuracy on the validation (trial) set.

@@

    \caption{Ablation study results.}
    \vspace{-0.4cm}
    \caption*{
        \footnotesize Accuracies were obtained from the
        validation dataset. Each model was trained with the same random seed and
        hyperparameters, save for the one listed. ``No emoji'' is the same model
        trained on the training dataset with no emoji, ``No ELMo'' corresponds
        to having switched the ELMo word encoding layer with a simple
        pre-trained GloVe embedding lookup table, and ``Concat Pooling''
        obtained sentence representations by using the pooling method described
        by \citet{howard2018universal}. ``LSTM hidden'' corresponds to the
        hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the
        part-of-speech embeddings, and ``SGD optim lr'' to the learning rate
        used while optimizing with the schedule described by \citet{conneau2017supervised}.}

@@

Finally, not all emoji are beneficial for this task. When removing
\texttt{sweat\_smile} and \texttt{confused} accuracy increased, probably because
they represent emotions other than the ones being predicted.

%From observing Table \ref{table:emoji_label} it seems likely that tweets
%containing \texttt{mask} would be highly predictive of the \texttt{disgust}
%class, since more than $92\%$ of the tweets containing such emoji belong to
%that class, however it less clear why \texttt{rage}, and \texttt{cry}
%contribute similarly, despite being more evenly distributed among classes.

%This means that they are highly representative of the the class the tweet
%belongs to, and would be good choices for discriminating new tweets in case we
%wanted to expand the dataset.

%For example, by observing Table \ref{table:emoji_label} we can infer that
%\texttt{mask}, \texttt{rage}, and \texttt{cry}   \texttt{disgust},
%\texttt{rage}, and \texttt{sad} classes respectively.

@@

\section{Acknowledgements}%
\label{sec:acknowledgements}

We thank the anonymous reviewers for their reviews and suggestions. The first
author is partially supported by the Japanese Government MEXT Scholarship.
